---
title: "Project 2 Exploratory data analysis using a supplied dataset"
author: "J. Hamski"
date: "March 7, 2015"
output: html_document
---
***

#Data Preparation

The data for this project was only provided in the PDF. So, I performed the following steps to prepare it for loading into R.
1. Created a screenshot (data_table.png) of the table only.
2. Uploaded the screenshot image to Online OCR http://www.onlineocr.net/
3. Converted data table using Online OCR to a MS Excel .xlsx file
4. Spot-checked the data and fixed one error where the OCR read an "8" as "a"
5. Saved the data as OCR_data_table.csv

This method worked reasonably well for this small image. A privately hosted or desktop solution would be better if the data needs to be kept private. Additional QA/QC would be appropriate for a larger dataset. 

> Packages used: datasets, ggplot2, dplyr

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```
```{r}

input <- read.csv("OCR_data_table.csv", skip = 1)

#below is dummy data, still need to format input
X1 <- c(10.0, 8.0, 13.0)
Y1 <- c(8.04, 6.95, 7.58)
X2 <- c(10.0, 8.0, 13.0)
Y2 <- c(9.14, 8.14, 8.74)
X3 <- c(10.0, 8.0, 13.0)
Y3 <- c(7.46, 6.77, 12.74)
X4 <- c(8.0, 8.0, 8.0)
Y4 <- c(6.58, 5.76, 7.71)

DF <-data.frame(X1, Y1, X2, Y2, X3, Y3, X4, Y4)
```

## Summary Statistics

```{r}
summary(DF)
```


## Graphic Exploration 
```{r, warning=FALSE}
library(ggplot2)

#Jim's House Theme

#JHT <- theme()
```

Basic Scatterplot for Category 1
```{r}
ggplot(data = DF, aes(x = X1, y = Y1)) + geom_point() + ggplot(data = DF, aes(x = X2, y = Y2)) + geom_point()
```

Basic Scatterplot for Category 1
```{r}
ggplot(data = DF, aes(x = X2, y = Y2)) + geom_point()
```

Basic Scatterplot for Category 1
```{r}
ggplot(data = DF, aes(x = X3, y = Y3)) + geom_point()
```
