---
title: "Project 2 - Exploratory data analysis using a supplied dataset"
author: "J. Hamski"
date: "March 7, 2015"
output: html_document
---
***
##Assignment Requirements
1. Perform an exploratory data analysis (EDA) on the dataset and draw conclusions
2. Use ggplot2
3. Deliver report in Markdown on RPubs  
  
> Packages used: dplyr, tidyr, ggplot2, RColorBrewer, grid, scales
  
##EDA Resources

From the [EDA Wikipedia Page](http://en.wikipedia.org/wiki/Exploratory_data_analysis), the objectives of EDA are:  
1. Suggest hypotheses about the causes of observed phenomena  
2. Assess assumptions on which statistical inference will be based  
3. Support the selection of appropriate statistical tools and techniques  
4. Provide a basis for further data collection through surveys or experiments  

I relied on [*Experimental Design for Behavioral and Social Sciences, Ch. 4, EDA (Seltman, H; 2013)*](http://www.stat.cmu.edu/~hseltman/309/Book/) for guidance on the statistical underpinnings of EDA. I referenced [Tutorial 3: Graphics and Exploratory Data Analysis in R (Piennar, J. and Miller, T.; 2013)](http://bio.fsu.edu/miller/docs/Tutorials/Tutorial3_Graphics.pdf) and [Introduction to R and Exploratory data analysis (Simpson, G; 2006)](http://cc.oulu.fi/~jarioksa/opetus/metodi/eda.pdf) for R-specific EDA techniques. 

##Data Preparation

The data for this project were provided in a PDF. So, I performed the following steps to prepare it for loading into R.  
1. Created a screenshot (data_table.png) of the table only.  
2. Uploaded the screenshot image to Online OCR http://www.onlineocr.net/  
3. Converted data table using Online OCR to a MS Excel .xlsx file  
4. Spot-checked the data and fixed one error where the OCR read an "8" as "a"  
5. Saved the data as 'OCR_data_table.csv'  

This method worked well for this small image. A privately hosted or desktop OCR solution would be better if the data needs to be kept private. Additional QA/QC would be required for a larger dataset. 

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
```
```{r}
input <- read.csv("OCR_data_table.csv", skip = 1)
DF.1 <-data.frame(input[1:2])
colnames(DF.1) <-c("x", "y")
DF.2 <-data.frame(input[3:4])
colnames(DF.2) <-c("x", "y")
DF.3 <-data.frame(input[5:6])
colnames(DF.3) <-c("x", "y")
DF.4 <-data.frame(input[7:8])
colnames(DF.4) <-c("x", "y")

DF.all<- cbind(DF.1, DF.2, DF.3, DF.4)
```
***
#SECTION 1 - Bivariate Analysis
***
##Dataframe 1 - "I"
###Summary Statistics and Non-graphical EDA

Both parameters in DF.1 are numeric variables. Their summary statistics are:
```{r}
summary(DF.1)
```
X and Y have a similar range, with a lower bound of approximately 4 and an upper bound of approximately 14 and 10, respectively. The central tendency of the two parameters diverge by approximately 1.5 units with x being the larger mean (9 and 7.5, respectively).

```{r}
sd(DF.1$x)
sd(DF.1$y)
```
The spread of the two parameters is 

## Graphical EDA 
This EDA uses my own "house theme" that I developed. The default graphics for ggplot2 are stellar, but by incorporating your own theme you can make it easier to identify and place your own work. See Appendix 1 - Jim's House Theme for more information
```{r, warning=FALSE}
library(ggplot2)
library(RColorBrewer)
library(scales)
library(grid)

```


Basic Scatterplot for Category 1
```{r}
#ggplot(data = DF.1, aes(x = x, y = y)) + geom_point() + JHT()
```

##Dataframe 2 - "II"
###Summary Statistics and Non-graphical EDA
## Graphical EDA 




***
SECTION 2 MULTIVARIATE ANALYSIS  
***

##Appendix 1 - Jim's House Theme

```{r}
#Jim's House Theme
JHT <- function(){
# Generate the colors for the chart procedurally with RColorBrewer
  palette <- brewer.pal("Greys", n=9)
  color.background = palette[1]
  color.grid.major = palette[3]
  color.axis.text = palette[6]
  color.axis.title = palette[9]
  color.title = palette[9]
  
  # Begin construction of chart
  theme_bw(base_size=9) +
    
    # Set the entire chart region to a light gray color
    theme(panel.background=element_rect(fill=color.background, color=color.background)) +
    theme(plot.background=element_rect(fill=color.background, color=color.background)) +
    theme(panel.border=element_rect(color=color.background)) +
    
    # Format the grid
    theme(panel.grid.major=element_line(color=color.grid.major,size=.25, linetype = "longdash")) + 
    theme(axis.line = element_line(color=color.grid.major, size = .5))+
    theme(panel.grid.minor=element_blank()) +
    theme(axis.ticks=element_blank()) +
    
    # Format the legend, but hide by default
    theme(legend.position="none") +
    theme(legend.background = element_rect(fill=color.background)) +
    theme(legend.text = element_text(size=7,color=color.axis.title)) +
    
    # Set title and axis labels, and format these and tick marks
    theme(plot.title=element_text(color=color.title, size=10, vjust=1.25)) +
    theme(axis.text.x=element_text(size=12,color=color.axis.text)) +
    theme(axis.text.y=element_text(size=12,color=color.axis.text)) +
    theme(axis.title.x=element_text(size=14,color=color.axis.title)) +
    theme(axis.title.y=element_text(size=14,color=color.axis.title)) +
    
    # Plot margins
    theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}
```